Geometric variational problems

### 2020-11-05 NSF2020/Description/specificaim1_ver2.tex
mod SA1v2

\subsection{Specific Aim 1: Collective elastic and hydrodynamic properties of amphiphilic particles}
\label{subsec:specific_aim_1}

%\begin{figure}
%\begin{center}
%\includegraphics[width=0.9\textwidth]{figures/SA1_fig1.pdf}
%\end{center}
%\caption{\footnotesize (A--C) The splay ($\Div \mathbf{d}$), 
%saddle splay ($\det \mathsf{D}$) and twist ($\Curl \mathbf{d}$) elastic distortions of 
%a monolayer. (D) The monolayer neutral surface $\Sigma$,  
%director $\mathbf{d}$ and unit normal $\mathbf{n}$ in local coordinates.
%(E--F) Lipids are able to tilt from  the surface normal, and stretch.
%(H--J) Intermediates of membrane fusion and (K) experimental image of a stalk \cite{Aeffner2012}. }
%\label{fig:distortions}
%\end{figure}

@@

@@

The elastic theory assumes an energy density that is quadratic in the Green-Lagrange strain tensor for $x$. 
The chains stretch and compress to satisfy incompressibility.
Symmetries with respect to mirror reflection and in-plane rotation, the internal structure assumptions and incompressibility
lead to an elastic surface energy over the dividing surface $\Sigma$. 
This energy decomposes into four, fundamental and independent deformations:
splay ($\Div n$), twist ($\Curl n$), saddle splay ($\det \nabla n$) and tilt $T$:
\begin{equation}
\label{ansatz3}
\begin{aligned}
\int_{\Sigma} 
  \tfrac{1}{2}\KB\left[ \left( \Div n + k_0\right)^2 - k_0^2\right] 
+ \tfrac{1}{2}\KT (\Curl n)^2 + \KG  \det \nabla n + \KTH |T|^2 \,dA.
\end{aligned}
\end{equation}
Here $\Div n$, $\Curl n$ and $\nabla n$ refer to the surface divergence, surface curl and surface gradient operators
respectively (Figure \ref{fig:distortions}A--C).
The deformations come with elastic coefficients: the \emph{bending modulus} $\KB$, \emph{twist modulus} $\KT$ 
and \emph{saddle-splay modulus} $\KG$ and \emph{tilt modulus} $\KTH$
HK define the tilt vector $T = n/(N\cdot n) - N$ where $N$ is the unit surface normal.

@@

An interesting and challenging feature of \eqref{ansatz3} is that
the elastic energy couples director gradients to surface geometry through the tilt vector field.
In biological contexts, tilt appears at lipid domain boundaries, the edge of nanopores, and at the boundary of membrane inclusions \cite{https://doi.org/10.1103/PhysRevE.102.042406}, and the bilayer energy is the sum monolayer energies. 
Under spatial scales much larger than the membrane thickness, membrane energy is well-characterized by
the Canham-Helfrich energy used throughout the fluid-structure literature \cite{}.
The Canham-Helfrich energy is actually a special case of \eqref{ansatz3} obtained by setting $n =  \pm N$ (the $\pm$ depending
on orientation) and collapsing both monolayers onto the membrane midplane. There,
splay $\Div n$ equal to twice the mean curvature, saddle splay $\det \nabla n$ equal to the Gauss curvature,
and both twist and tilt equaling zero.

@@

Recently, there has been a revival in interest in the HK theory. The quadratic assumption
for the elasticity energy density has caused researchers to question the validity of \eqref{ansatz}
for situations involving large curavtures \cite{PRL}. Zimmerberg and Akimov \cite{https://doi.org/10.1039/C9SM02079A}
showed that energies derived from molecular dynamics and elasticity agree even in the case of large deformations.
Part of this proposal will contribute to clarifying some of the mathematical
assumptions that went into the HK theory.

\subsubsection{Evidence for elastic properties}
\label{sssec:evidence}
We tested for elastic properties by subjecting two-dimensional bilayer morphologies
to external loads \cite{Fu19}, using realistic
values for phospholipid length \cite{Boal},
the screening length $\rho = 2.5$ nm \cite{Eriksson1989,Lin2005,Parsegian,Israelachvili80,TerziDeserno17}
and  interfacial tension $\gamma=4.1$ pN nm$^{-1}$  \cite{GarciaSaez, KUZMIN2005, Petelska2012}.
Throughout the proposal, \kBT\; = 4.11$\times 10^{-21}$ J.
We first considered a planar bilayer subject to a uniform vertical load on the particle centers (Figure \ref{fig:bend}). 
The bilayer is clamped and horizontal at one end and the restoring force in the free part 
of the bilayer opposes the load. Twist and saddle splay are both zero, singling out splay
as the only distortion.
% In this situation, the directors are also parallel to the unit surface
%normal. As such $\Div \mathbf{d} = -2H$ where $H$ is the mean curvature of the surface. 
Since deformations are small \eqref{ansatz3}, it is possible to solve the bilayer loading in closed form. 
This analytical solution (red curve) basically overlaps the midplane of the particle based solution (Figure \ref{fig:bend}A).

Experimental measurements have accurately determined 
the bending modulus for several lipid types, with 
typical values lying around 10 \kBT\; \cite{Naetal15,VeBrPa15,NAGLE2000159,PhysRevLett.113.248102}.
By comparing with the exact solution, we derived a bending modulus
in the range 8.51 -- 13.54 \kBT, which is in 
excellent agreement with the experimentally derived value, especially considering how few parameters the HAP 
formalism involves.

@@

\subsubsection{What work will actually be done?}
% Proposers should address what they want to do,
% why they want to do it,
% how they plan to do it,
% how they will know if they succeed,
% and what benefits could accrue if the project is successful.

We will study the elastic properties of particle-based bilayers using static and hydrodynamic, large particle-number simulation. 
The goal is to connect the deformations of particle bilayers to the deformations of HK theory, and in particular map the parameters of the
particle-based system onto known, and experimentally measured elastic moduli.

The successful outcomes of our preliminary tests suggest that the HAP particle configurations 
do behave like an elastic material. To validate our conjecture, we must consider 
ways to isolate or partially isolate the remaining distortions.

\subsubsection{Hydrodynamic simulations}

\subsubsection{Static simulations}
Particle-based bilayers form because of the balance between hydrophobic attraction and excluded volume repulsion.
As a result, a precise way to measure a certain elastic properties is to set up an externally loaded system,
and then isolate the desired deformation by solving for the energy minimizer. The following give technical details on how we
draw these connections in practice.
In summary, the elastic moduli determine how lipid monolayers and bilayers mechanically behave.

\begin{description}
\item[Bending modulus] A planar bilayer that is deformed into the shape of a circular arc stores bending energy in the form of splay.
  Bending isolates the splay deformation $\Div{n}$ because the twist and saddle-splay deformations are zero in two-dimensional bilayers,
  and both MD and continuum mechanical simulations show that tilt is more or less zero due to lack of inclusions.
  Moreover, the splay in the two leaflets of an arc approximately cancel when added, due to the smaller head to tail ratio in the inner leaflet and
  larger head to tail ratio in the outer lealet, and so we can neglect the spontaneous curvature.

  Using steepest gradient descent, the simulation tracks the relaxation of the energies $E$ of a sequence of arc-shaped bilayers tending toward and the equilibrium, planar bilayer.
  We reconstruct the monolayer dividing surface (a curve $C$) and directors by interpolating the particle centers and orientations. Using spectral differentiation,
  we can compute the curvature $\kappa$ of $C$ and the squared curvature energy $H = \int_C \kappa^2$. Since $\Div{n} = -\kappa$ in this setup, we can plot
  the particle based values $\Delta E = E - E_0$ as a function of $\Delta H = H - H_0$ where $E_0$ and $H_0$ are the respective equilibrium values.

  Figure \ref{} shows the appoximately linear relationship between $\Delta E$ and $\Delta H$.
  The slope of the dashed line gives the bending modulus $\KB = 10$ kT. 
  The simulation is for a realistic phospholipid length \cite{Boal},
  the screening length $\rho = 2.5$ nm \cite{Eriksson1989,Lin2005,Parsegian,Israelachvili80,TerziDeserno17}
  and  interfacial tension $\gamma=4.1$ pN nm$^{-1}$  \cite{GarciaSaez, KUZMIN2005, Petelska2012}.

  The parameters of our hydrophobic attraction formalism map accurately onto the experimentally determined moduli.  
  The value $\KB = 10$ kT is a reasonable bending modulus for many lipid species \cite{}, and it depends linearly on
  the interfacial tension parameter $\gamma$ of our model. The $\gamma = 4.1$ pN nm$^{-1}$ is
  about a third of the surface tension of a typical hydrocarbon-water interface. M. Jackson points out, however, that the smaller interfacial tension parameter is actually the correct value 
  to describe the free energy for placing molecular lipid into contact with water \cite{}.
  So not only is the HARP model tunable to experimental parameters, but the results coincide with other properties from physical chemistry.

\item[Tilt]
\item[Spontaneous curvature]
\item[Twist and saddle-splay]
  We measure the elastic response to twist by following the relaxation
  dynamics of an initially twisted membrane. Tilt is a fully three-dimensional deformation
  and so the reference membrane is a pancake-shaped, planar bilayer with diamater
  much larger than bilayer thickness.  In the initial configuration, the 
  the directors behave to leading order as $n = (0, mx_1, \sqrt{1 - m^2x_1^2})$
  where $x_1$ is the twisting axis. At the continuum level, 
  the surface gradient $\nabla n = \begin{pmatrix}0 & 0\\ m & 0\end{pmatrix}$ that $\Div n = \mathrm{tr}(\nabla n) = 0$,
  $\det(\nabla n) = 0$ and $\Curl d = (\nabla u)_{12} - (\nabla u)_{21} = m$,
  so that twist is isolated and the continuum elastic energy equals $\KA m^2 A + O(1)$ where
  $A \gg 1$ is the membrane area.

  The initial bilayer consi
The twist deformation, for example, has a modulus in the range of 1 to 2 \kBT\; 
as evaluated by molecular dynamics \cite{LeVeWa14}. 
But twist is mathematically zero in our two-dimensional monolayers,
and so we require a computationally efficient means for evaluating the 
boundary integral equations in a collection of three-dimensional particles. 
Specific Aim 2 discusses some of the outstanding implementation issues.

Our eventual three-dimensional calculations must also consider the saddle-splay distortion,
and the calculations may help resolve some well-known issues in the theory \cite{TerziDeserno17}. 
Theoretical analysis of lipid phase transitions predict a negative saddle-splay modulus around $-8$ \kBT\;
\cite{SIEGEL2004366,SIEGEL20085200}.
This value, however, 
predicts an energy barrier for monolayer fusion on the order of 200  \kBT, which runs counter to 
the value 30 \kBT\; estimated in experiments \cite{RyKlYaCo16,FrRoPi17,Tran7106}.  
We expect the particle-based approach can shed light on this inconsistency.   
\item[Stretching]
We also considered the stretching deformation \cite{Fu19}. 
Stretching occurs whenever there is an excess monolayer area,
and it is energetically costly since it exposes hydrocarbon tails to water. 
%Monolayers are not, strictly speaking, area incompressible,
%and 
Manipulation experiments give a monolayer area modulus in the range 
30 -- 40 \kBT\; nm$^{-2}$ \cite{Nagle17, Nagle17-2}. 
To measure an area modulus, we stretched a circular vesicle, modeling the cross-section 
of a bilayer tube (Figure \ref{fig:bend}B \& C).
%We then stretched the vesicle beyond and beneath equilibrium radius and recorded the change in energy. 
This procedure gave a monolayer area modulus 
34 $\pm$ 2 \kBT \;nm$^{-2}$, independently of the vesicle size, which
is also remarkably close to the experimental value.

\item[Fluctuation spectrum]
\item[Revision of HK Theory]
  It may come as a surprise that the field of membrane continuum mechanics still lacks concensus as to
  whether \eqref{ansatz3} contains a complete list of consistent deformations.
  The PI Ryham showed, for example, that the tilt energy density is interchangagle, up to second order, with
  the simplified term $|n \times N|^2$ \cite{RyKlYaCo16}.
  Recently, \cite{10.1063/1.4990404} derived a tilt curvature term that Hamm and Kozlov had neglected from their analysis \cite{HaKo2000}.
  Later, \cite{https://doi.org/10.1039/C9SM02079A} 
  and \cite{10.1103/PhysRevE.102.042406} independently identified an inconsitency in the transversal tilt invariance
  assumption of Desrno's earlier work \cite{10.1063/1.4990404}.

  We suspect the reason why, after twenty years, the field has yet to reach a concensus 
  lies in the asymptotic expansions used to derive the elastic energy density. 
  Past researchers expressed the strain tensor in
  terms of the coordinate frame $\{\tau_1, \tau_2, N\}$ with surface tangent vectors $\tau_i$ and surface normal $N$.  
  In this frame, there is a coupling between the gradients of all three terms $x_0,$ $\zeta$ and $n$ of the deformation \eqref{LMdeformation}.
  The differential geometric arguments needed to determine which terms of the expansion to retain in the expansion are subtle,
  and have caused \cite{10.1103/PhysRevE.102.042406} to derive an energy density with more than ten elastic parameters, for example.

  However, expanding the strain tensor in terms of the coordinate frame $\{e_1, e_2, n\}$ with $e_i$ perpendicular to $n$
  decouples a good portion of the gradient terms, leading to simplified expressions. For example,
  in this coordinate frame, the incompressility condition leads to an exact relationship 
  \begin{equation}
    \label{elongation}
    (1 + \Div n \zeta + \det \nabla n \zeta^2) \zeta' = 1
  \end{equation}
  for lipid elongation. \cite{HaKo2000} seem to have been aware of \eqref{elongation} for the special case of zero tilt,
  but the works \cite{HaKo2000, https://doi.org/10.1039/C9SM02079A, 10.1063/1.4990404} use an approximate form of \eqref{elongation}
  and \cite{10.1103/PhysRevE.102.042406} reports a form containing an erroneous $|T|^2$ in the $\zeta^0$ coefficient.

Our investigations will leave open the possibility that the free energy of monolayers and bilayers is non-local. 
We have shown that hydrophic attraction is a non-additive force, meaning the force of attraction between two
particles is dependent on the position a third or more particles. Conceptually, this would arise in continuum theory if two monolayer surfaces
had different elastic energy density at that point despite having locally identical curvatures and directors.

\end{description}

>>>>

\subsection{Specific Aim 1: Collective elastic properties of amphiphilic particles}
\label{subsec:specific_aim_1}

%\begin{figure}
%\begin{center}
%\includegraphics[width=0.9\textwidth]{figures/SA1_fig1.pdf}
%\end{center}
%\caption{\footnotesize (A--C) The splay ($\Div \mathbf{d}$), 
%saddle splay ($\det \mathsf{D}$) and twist ($\Curl \mathbf{d}$) elastic distortions of 
%a monolayer. (D) The monolayer neutral surface $\Sigma$,  
%director $\mathbf{d}$ and unit normal $\mathbf{n}$ in local coordinates.
%(E--F) Lipids are able to tilt from  the surface normal, and stretch.
%(H--J) Intermediates of membrane fusion and (K) experimental image of a stalk \cite{Aeffner2012}. }
%\label{fig:distortions}
%\end{figure}
The goal of Specific Aim 1 is characterize the material properties of many-body, self-assembled amphiphiles.
For amphiphiles assembled into bilayers, these properties are described by membrane continuum mechanics.
The end-goal is to be able to map the parameters of the particle-based model onto the elastic moduli from continuum theory.
If we achieve this goal, then simulators can use the hydrophobic attraction force calculations
to model bilayers with specific composition. These calculations have provably less computational complexity than
those of molecular dynamics simulations and possess the molecular granularity lacking from continuum models.

@@

Recently, there has been a revival in interest in the HK theory. The quadratic assumption
for the elasticity energy density has caused researchers to question the validity of \eqref{ansatz}
for situations involving large curavtures \cite{PRL, https://doi.org/10.1016/j.bbamem.2016.02.003}. Zimmerberg and Akimov \cite{https://doi.org/10.1039/C9SM02079A}
showed that energies derived from molecular dynamics and elasticity agree even in the case of large deformations.
Part of this proposal will contribute to clarifying some of the mathematical
assumptions that went into the HK theory.

@@

The elastic theory assumes an energy density that is quadratic in the Green-Lagrange strain tensor for $x$. 
The chains stretch and compress to satisfy incompressibility.
Symmetries with respect to mirror reflection and in-plane rotation, the internal structure assumptions and incompressibility
lead to an elastic surface energy over the dividing surface $\Sigma$. 
This energy decomposes into four, fundamental and independent deformations:
splay ($\Div n$), twist ($\Curl n$), saddle splay ($\det \nabla n$) and tilt $T$:
\begin{equation}
\label{ansatz3}
\begin{aligned}
\int_{\Sigma} 
  \tfrac{1}{2}\KB\left[ \left( \Div n + k_0\right)^2 - k_0^2\right] 
+ \tfrac{1}{2}\KT (\Curl n)^2 + \KG  \det \nabla n + \KTH |T|^2 \,dA.
\end{aligned}
\end{equation}
Here $\Div n$, $\Curl n$ and $\nabla n$ refer to the surface divergence, surface curl and surface gradient operators
respectively (Figure \ref{fig:distortions}A--C).
The deformations come with elastic coefficients: the bending modulus $\KB$, twist modulus $\KT$ 
and saddle-splay modulus $\KG$ and tilt modulus $\KTH$
HK define the tilt vector $T = n/(N\cdot n) - N$ where $N$ is the unit surface normal.

@@

An interesting and challenging feature of \eqref{ansatz3} is that
the elastic energy couples director gradients to surface geometry through the tilt vector field.
In biological contexts, tilt appears at lipid domain boundaries, the edge of nanopores, and at the boundary of membrane inclusions \cite{https://doi.org/10.1103/PhysRevE.102.042406}, and the bilayer energy is the sum monolayer energies. 
Under spatial scales much larger than the membrane thickness, membrane energy is well-characterized by
the Canham-Helfrich energy used throughout the fluid-structure literature \cite{}.
The Canham-Helfrich energy is actually a special case of \eqref{ansatz3} obtained by setting $n =  \pm N$ (the $\pm$ depending
on orientation) and collapsing both monolayers onto the membrane midplane. There,
splay $\Div n$ equal to twice the mean curvature, saddle splay $\det \nabla n$ equal to the Gauss curvature,
and both twist and tilt are equal to zero.

% Proposers should address what they want to do,
% why they want to do it,
% how they plan to do it,
% how they will know if they succeed,
% and what benefits could accrue if the project is successful.
\begin{center}
Preliminary results
\end{center}
Our preliminary tested for elastic properties by subjecting bilayers to external loads \cite{Fu19}.
We used realistic values for the phospholipid length $a$ \cite{Boal},
screening length $\rho = 2.5$ nm \cite{Eriksson1989,Lin2005,Parsegian,Israelachvili80,TerziDeserno17}
and  interfacial tension $\gamma=4.1$ pN nm$^{-1}$  \cite{GarciaSaez, KUZMIN2005, Petelska2012, Jackson}
model parameters.
The repulsion strength, repulsion length scale, particle shape (e.g. disks, ellipses), particle diameter and the hydrophobic boundary condition \eqref{}
are additional parameters defining the coarse-grained representation.

To derive a bending modulus, we first considered a planar bilayer subject to a uniform vertical load on the particle centers (Figure \ref{fig:bend}). 
Using a clamped condition at one end, the restoring force in the free part of the bilayer opposes the load.
For small deformations, it is possible to solve the bilayer loading in closed form, and 
this analytical solution (red curve) basically overlaps the midplane of the particle based solution (Figure \ref{fig:bend}A).
We thereby derived a bending modulus $\KB$ in the range 8.51 -- 13.54 \kBT. Throughout the proposal, \kBT\; = 4.11$\times 10^{-21}$ J.
Experimental measurements have accurately determined the bending modulus for several lipid types, with 
typical values lying around 10 \kBT\; \cite{Naetal15,VeBrPa15,NAGLE2000159,PhysRevLett.113.248102}.

We also considered the stretching and tilt deformations \cite{Fu19}. 
Stretching occurs whenever there is an excess monolayer area,
and it is energetically costly since it exposes hydrocarbon tails to water.
To measure an area modulus, we stretched a circular vesicle, modeling the cross-section 
of a bilayer tube (Figure \ref{fig:bend}B \& C).
This procedure gave a monolayer area modulus 
34 $\pm$ 2 \kBT \;nm$^{-2}$. Manipulation experiments give a monolayer area modulus in the range 
30 -- 40 \kBT\; nm$^{-2}$ \cite{Nagle17, Nagle17-2}.

Finally, we considered the tilt deformation by applying a fixed tilt angle boundary condition. 
This boundary condition produces a tilt field that decays with the length scale $l = \sqrt{\KB/\KTH}$ 
Fitting to this field, our particle based simulation gave $\lambda = 1.2$ nm, which matches
the value $\KTH \approx 10$ \kBT \; nm$^{-2}$ reported in the literature \cite{https://doi.org/10.1063/1.4932971}.

Our preliminary work is encouraging because in all cases, the bending, stretch, and tilt elastic moduli for the collection of amphiphiles
were in excellent agreement with the experimental values. This fact is especially remarkable 
since we assigned the main model parameters $\rho$ and $\gamma$ reasonable physical values, rather than using them as tuning parameter. 
Nevertheless, the characterization of material properties is incomplete. For one, the preliminary work involved relatively few particles, and we need to
perform convergence studies with large particle-number systems to arrive at conclusive results. Secondly, we have yet to establish a functional
relationship between the model parameters and elastic moduli. Finally, to have relevence in physics and biology
we must incorporate the remaining twist and saddle splay deformations.
The following outlines the simulations we will use to characterize material properties.

\begin{center}
Anticipated model simulations
\end{center}
To characterize material properties, we compare the energies and forces generated by
a collection of amphiphilic particles against the energies and forces of a continuum bilayer.
The comparison enables us to understand how the model parameters map onto elastic parameters.

In the preliminary work, bilayer shapes minimized energy with respect to an external force. Energy minimality is somewhat unrealistic because
biological membranes are constantly in motion. Tracking the evolution of a bilayer is a more robust procedure that does not require minimality.
Here we start with a particle bilayer in a non-equilibrium shape, and then evolve the particle bilayer according to steepest gradient descent with respect to the
HARP functional. Gradient descent stabilizes the evolving bilayer shape and gives a monotonically decreasing energy. This gives a relatively large
data set of energy $E$ versus particle configuration. At the same time, we reconstruct an evolving monolayer dividing surface $\Sigma$ and director field $n$ by
interpolating the particle centers and orientations. Using \eqref{ansatz}, we calculate a continuum energy $H$ from the interpolated
bilayer shape. Setting $\Delta E = E - E_0$ and $\Delta H = H - H_0$ normalizes the energies to their respective equilibrium values $E_0$ and $H_0$.
In certain instances, the bilayer evolution involves only one of the four fundamental deformations.
When this occurs, the elastic moduli appear as the slope of the linear regression between $\Delta H$ and $\Delta E$.

Bending is the energetically most consequential deformations,
and so we revisit estimating the bending modulus parameter. 
In the case of the bending deformation, we deform the midplane of a planar bilayer into the shape of a circular arc.
This arc stores bending energy in the form of splay $\Div{n}$, while twist, saddle-splay and tilt deformations are zero.
The splay in the two leaflets approximately cancel when added, making the contribution of spontaneous curvature negligible.
Since $\Div{n} = -\kappa$ in this setup, we can compute the curvature squared curvature energy 
$H = \int_C \kappa^2$ of the cross-section $C$, a simple closed curve. Figure \ref{} shows the appoximately linear
relationship between $\Delta E$ and $\Delta H$ with $\KB$ as the slope.

The twist deformation is a fully three-dimensional deformation. 
(Specific Aim 3 address outstanding implementation issues like three-dimensional boundary integral solvers.) 
The twist deformation is closely related to the assumption of lateral fluidity in membranes and so researchers 
have sometimes assumed $\KT = 0$ \cite{HaKo2000, 10.1063/1.4990404, https://doi.org/10.1039/C9SM02079A, 10.1103/PhysRevE.102.042406}, 
while molecular dynamics investigations find a twist modulus in the range of 1 to 2 \kBT \cite{LeVeWa14}.
To measure the elastic response,  we initially twist the bilayer along an $x_1$-coordinate axis.
In this configuration, the directors behave at leading order as $n = (0, mx_1, \sqrt{1 - m^2x_1^2})$,
yielding a surface gradient $\nabla n = \begin{pmatrix}0 & 0\\ m & 0\end{pmatrix}$. 
Here $\Div n = \mathrm{tr}(\nabla n) = 0$, $\det(\nabla n) = 0$ and $\Curl d = (\nabla u)_{12} - (\nabla u)_{21} = m$,
so that the continuum elastic energy equals $\tfrac{1}{2}\KT m^2 A + O(1)$ for a membrane area $A \gg 1$. (Our particle collections
conserve membrane area, unless they lie in a strong background flow.) 
This enables us to plot the HARP energy values derived from steepest gradient descent as a function of $m$, and finally
yields a value for $\KA$.

We will also determine the effective saddle splay modulus $\KG$. 
Theoretical analysis of lipid phase transitions predict a negative saddle-splay modulus around $-8$ \kBT\;
\cite{SIEGEL2004366,SIEGEL20085200}. Researchers point out, however, that this theoretical value requires a larger energy 
barrier for monolayer fusion than is found by experiment \cite{FrRoPi17,Tran7106,TerziDeserno17}. 
The field requires further theoretical work like the what we are proposing to more precisely account for saddle splay.

The gradient descent technique is ineffective for measuring $\KG$ because
the saddle splay energy is largely invariant under shape changes.
In the simplified Canham-Helfrich formulation, saddle splay energy is an indicator of topological transitions, thanks to the  Gauss-Bonnet theorem
\cite{TerziDeserno17}.
PI-Ryham showed that saddle splay acts as a topological indicator even in the presence of nonzero tilt \cite{RyKlYaCo16}.

To evaluate $\KG$, we will combine the string method from PI-Ryham's work on fusion \cite{RyKlYaCo16} with
the present particle simulations. The string method is a numerical scheme that finds
least energy pathways separating energy basins \cite{simple_string}.
Using the string method, we will evaluate the transition energies between the intact pancake shape,
and one with a pore. Pore formation in a single bilayer decreases the integral of saddle splay from $4\pi$
to $0$. We can, in principle, detect this drop in energy becuase we have already accounted for the other deformations.

We will also analyze $\KG$ using a helical geometry. In a helicoid with inclination $m$, the director
gradient is $\nabla n = \begin{pmatrix} 0 & m(m^2+x_1^2)^{-1} \\ m(m^2+x_1^2)^{-1} & 0\end{pmatrix}$ where $x_1$ is the distance to the central
axis. We observe that saddle splay, twist and tilt are zeros, and $\det \nabla n = m^2(m^2+x_1^2)^{-2}$.
In the simulations, the particles forming the helicoid will be confined to two, concentric cylinders. The
continuum energy $\tfrac{1}{2}\KG h(m x_1(m^2+x_1^2)^{-1} + \tan^{-1}(x_1/m)) |_{r_1}^{r_2}$ where $h$, $r_1$ and $r_2$
are the cylinder heights, inner cylinder radius, and outer cylinder radius, respectively.

@@

 Perhaps surprisingly, the field of membrane continuum mechanics still lacks consensus as to
  whether \eqref{ansatz3} contains a complete list of consistent deformations.
  The PI Ryham showed, for example, that the tilt vector $T = n/n\cdot N - N$ leads to unphysical cusps when constraining membrane thickness
 \cite{RyKlYaCo16}.
  Recently, \cite{10.1063/1.4990404} derived a tilt curvature term that was neglected from the HK analysis \cite{HaKo2000}.
  Later, \cite{https://doi.org/10.1039/C9SM02079A} 
  and \cite{10.1103/PhysRevE.102.042406} independently identified an inconsistency \cite{10.1063/1.4990404} arising
  from a transversal tilt invariance assumption.

  We suspect the reason why, after twenty years, the field has yet to reach a consensus 
  lies in the Taylor expansions used to derive the elastic energy density. 
  Past researchers expressed the strain tensor in
  terms of the coordinate frame $\{e_1, e_2, N\}$ with surface tangent vectors $e_i$ and surface normal $N$.  
  In this frame, there is a coupling between the gradients of all three terms $x_0,$ $\zeta$ and $n$ of the deformation \eqref{LMdeformation}.
  The recent paper \cite{10.1103/PhysRevE.102.042406} contains an energy density with more than ten elastic parameters, for example.
  Subtle differential geometric arguments are needed to evaluate the terms in the Taylor expansion.

  We will revisit the HK analysis to see if any inconsitencies can be resolved.
  As a main analytical innovation, we have found that expanding the strain tensor in
  terms of the coordinate frame $\{e'_1, e'_2, n\}$ with $e'_i$ perpendicular to $n$
  greatly simplifies the expansion because it decouples the gradient terms. As an illustration, the prominent works
  \cite{10.1063/1.4990404, 10.1103/PhysRevE.102.042406, HaKo2000, https://doi.org/10.1039/C9SM02079A} utilize an
  incompressility condition for the lipid elongation function $\zeta$ from \eqref{LMdeformation} that is only approximately true. 
  However, when expressed in the $\{e'_1, e'_2, n\}$ coordinate frame, the incompressility condition is exact
  and is given by a Steiner-type polynomial in terms of $\Div n$ and $\det \nabla n$ \cite{http://www.jstor.org/stable/1993504}.
  Aside from the special case of zero tilt, \cite{HaKo2000} an coworkers seem to have been unaware of this result.

Our investigations will leave open the possibility that the free energy of monolayers and bilayers is non-local. 
We have shown that hydrophic attraction is a non-additive force, meaning the force of attraction between two
particles is dependent on the position a third or more particles. Conceptually, this would arise in continuum theory if two monolayer surfaces
had a different elastic energy density despite having locally identical curvatures and directors.
The boundary integral formulation described in Specific Aim 3 easily accomodates additional linear PDE and
we will extend our study to include charge lipids and weak electrolytes \cite{DIMOVA}.

@@

Specific Aim 1 dealt with the collective material property of amphiphiles which does not involve the movement  of the solvent.
In Specific Aim 2, we include hydrodynamics effects. The reason for including hydrodynamic effects is that 
biologically relevant functions like fusion, fission, pore dynamics involve viscous dissipation of the aqueous environment. 
Moreover, including hydrodynamics adds relevance to our theory since most coarse-grained and molecular dynamics theories
include water either explicitly or implicitly in their models.
